Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU

### 2016-02-10 paper/paperSource.tex
automatic triggered changes

\emph{\textbf{One-body queries kernel}}.
Most of the query modules in Table~\ref{tb:queries} (except the SDH and RDF) are not that involved, only containing computations of fairly simple, one-body functions. These queries were coded as separate modules in our
system. Each of these modules take few attributes as input (e.g.,
atom selection, frames selection(for the autocorrelation functions),
number of atoms, etc.). The system pushes the data as it becomes
available onto these modules. The queries are being executed on the
selection and are put in a ``ready'' mode, awaiting the next frame's
data. First, the more basic queries, like total mass, are being
computed. The results of such queries are temporary stored (in main
memory) and are available for use anytime a more complex query needs
them.

\subsubsection{Two-body queries} In general, queries involving two-body functions are a
bit more complex and cannot provide the final result in a single
data read if a straightforward method is used for their computation.
However, in the proposed system we have incorporated a data
structure and an algorithm for the SDH (also RDF) from our previous
work that is suitable for push-based type of system. In this
subsection we give a brief description of the data structure (DM)
and the algorithm (DM-SDH) designed in~\cite{ytu:icde09, EDBT12} and
implemented in the system proposed in this paper. For more detailed
information, please refer to our previously published work on this
topic~\cite{ytu:icde09, EDBT12}

@@

\begin{figure}
 \centerline{ \includegraphics[width=0.65\columnwidth]{images/resolve.eps} }
 \caption{Computing minimum (i.e., length of solid lines) and
 maximum distance (i.e., length of dashed lines) range between two
cells}
 \label{fg:resolve}
\end{figure}

>>>>

\emph{\textbf{One-body queries kernel}}.
Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious. Figure \ref{fg:reduction} nicely represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel. Having an array of atom data structures, we reduce the array accumulating all the needed data from each atom's attributes. We don't expect much performance improvement from coalesced memory access, since the data structure for atom can eliminate any gain of spacial prefetch, though for specific types of queries it could be possible to aggregate certain attributes into merged arrays and take an advantage. In this case, the solution should be simple in order to keep it easily extendable for other simple one-body queries. We fetch all the atoms from GPU global memory, since we use references to atoms once. In this case, allocating shared memory and copying it over would complicate the implementation and possibly create additional overhead for one time reference usage.

\emph{\textbf{Two-body queries kernel}}.
In general, queries involving two-body functions are a
bit more complex and cannot provide the final result in a single
data read if a straightforward method is used for their computation.
However, in the proposed system we have incorporated a data
structure and an algorithm for the SDH (also RDF) from our previous
work that is suitable for push-based type of system. In this
subsection we give a brief description of the data structure (DM)
and the algorithm (DM-SDH) designed in~\cite{ytu:icde09, EDBT12} and
implemented in the system proposed in this paper. For more detailed
information, please refer to our previously published work on this
topic~\cite{ytu:icde09, EDBT12}

@@

\begin{figure}
 \centerline{ \includegraphics[width=0.65\columnwidth]{images/reduction} }
 \caption{Array reduction using coalesced memory access\cite{gpureduction}}
 \label{fg:reduction}
\end{figure}
